Globally-centered autocovariances in MCMC

### 2020-07-29 Paper/main.tex
Preliminary work on the introduction. Heading home.

Markov chain Monte Carlo (MCMC) is a sampling method used to obtain correlated realizations from a stochastic process. Due to this serial correlation in observations, a large number of MCMC samples are required to estimate features of the target distribution as compared to the independent ones. Therefore, it becomes imperative for an MCMC practitioner to have an idea about underlying dependence structure in the Markov chains. Autocovariance function (ACF) characterizes this dependence for a stationary Markov chain. Suppose $\{X_t; t \in \mathbb{Z}\}$ is a stochastic process with mean $\mu$ and follows second-order stationarity i.e. (1) $\mathbb{E}[X_t] = \mu$ is constant for all $t$, (2) $\mathbb{E}[X_t^2]$ is constant for all $t$, and (3) $\mathbb{E}[X_t X_{t+k}]$ is a function of $k$ only $\forall k,t \in \mathbb{Z}$. Then, the autocovariance function at lag-$k$ is defined as
%
\[
    \Gamma(k) = \Cov(X_t, X_{t+k}) = \mathbb{E}[(X_t - \mu)(X_{t+k} - \mu)^T].
\]

A popular sample autocovariance function is the empirical estimator. On a finite sample $X = (X_1, X_2, ..., X_n)^T$, the autocovariance at lag-$k$ is estimated as
%
\begin{equation} \label{eq:empirical_acf}
    \hat{\Gamma}(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}}(X_t - \bar{X})(X_{(t+\abs{k})} - \bar{X})^T
\end{equation}

where $\bar{X} = \sum_{t=1}^nX_t/n$. Calculation of exact ACF has been done for simple models like autoregressive (AR), moving average (MA) (\cite{quenouille1947notes}), and autoregressive-moving-average (ARMA) (\cite{box2015time}) models. However, for most of the complex sampling algorithms, we rely on sample estimators for ACF. Unfortunately, the estimator given in equation \ref{eq:empirical_acf} has been observed to have a some drawbacks. Firstly, it is not immune to outliers (see \cite{ma2000highly}) and secondly, it is usually non-informative in case of slowly mixing Markov chains for decently finite samples. We address the second issue in this paper by using multiple representative samples of the target distribution via parallel Markov chains. With convenient multi-core computation resources, it is possible to get multiple Markov chains for a target distribution. Our goal is to combine the information from all the Markov chains while calculating the autocovariances\\

 A commonly faced problem with the MCMC models is the slow mixing property. In case of multi modal target distributions, it has been observed that the Markov chains tend to get \textit{stuck} in a particular mode for a long time. As a consequence, the chain fails to explore the sample space. In such situations, single chain empirical ACF will underestimate the autocovariance giving misleading inferences.  We propose an autocovariance estimator called globally-centered autocovariance function (G-ACF) with a simple fix in centering which has been observed to show drastically better results in many cases. We will further study the impact of G-ACF in calculation of two important applications of ACF (1) asymptotic variance of Markov chain average and (2) effective sample size (ESS).\\

 MCMC ensures asymptotic convergence, however this is not practically very useful where the number of samples is finite. There are many criteria for terminating simulation like fixed-time rule, sequencial fixed-volume rule (see \cite{glynn1992asymptotic}), and using Gelman Rubin diagnostics (\cite{gelman1992inference}). The relative fixed-volume rule has been thoroughly discussed for MCMC by \cite{flegal2015relative} and \cite{gong2016practical}. We will use method of effective sample size (ESS) by \cite{vats2019multivariate} for terminating the simulations. The quality of estimation of $\Sigma$ is critical in estimating their ESS. We will successfully show that the G-SVE estimates the truth better and prevents the premature stopping of Markov chain.\\

In the next section we introduce the new sample autocovariance estimator and calculate its bias for finite samples. We observe that the bias term for G-ACF is similar to the bias of empirical ACF except for a few small order end-effect terms that vanish as $n \to \infty$. In section \ref{sec:variance_est}, we introduce the globally-centered SVE and give a proof of strong consistency, bias, and variance calculations. We will elaborate on the estimator for ESS using G-SVE in section \ref{sec:ess}. Section \ref{sec:examples} gives an experimental study on three different examples for the proposed estimators. All the important proofs are given in the Appendix.

@@

We demonstrate the striking difference in the ACF plots through mixture of Gaussian target density using a random-walk Metropolis-Hastings sampler. The target density has two well separated and non-equal modes. Two Markov chains are started close to each mode. It can be seen in Figure \ref{fig:gaussian-trace} that for the first $10^4$ samples, the chains do not communicate with the other mode. Whereas for a sample size of $10^5$, the chains have jumped modes and therefore, a single chain can give reliable information about the distribution. We wish to make two points here: firstly, for small sample size, we observe that ACF gives extremely misleading estimates of autocovariance whereas G-ACF is very close to reality (Figure \ref{subfig:gauss-acf-1e4}). Whereas, sharing the information about Monte Carlo average across chains leads to better estimation of autocovariances. Secondly, for a large sample size, the estimates from G-ACF as well as ACF are equivalent. Figure \ref{subfig:gauss-trace_1e5} depicts the scenario when convergence is achieved. This shows that G-ACF is at least as good as ACF; hence safe to use in any setting.

@@

 Suppose we have $m$ replications of Markov chains for same target distribution and $\{X_{st}; \; t \in \mathbb{Z}\}$ denote the $s^{th}$ Markov chain. Let the sample mean of $s^{th}$ Markov chain be denoted by $\bar{X}_s$ and the global mean by $\bar{\bar{X}} = \sum_{s = 1}^{m}\bar{X}_s/m$. The global mean provides a better estimate for the expectation value when the number of simulation per chain are not sufficient for the convergence to kick in. The globally-centered autocovariance function (G-ACF) for the $s^{th}$ Markov chain is given by:
%
\[
\hat{\Gamma}_{G,s}(k) = \dfrac{1}{n} \sum_{t=1}^{n-k}(X_{st}-\bar{\bar{X}})(X_{s(t+k)}-\bar{\bar{X}})^T
\]

If the starting points of these parallel Markov chains are sufficiently dispersed, G-ACF provides more realistic estimation of lag covariances. \cite{priestley1981spectral} shows that the empirical estimator in Equation \ref{eq:empirical_acf} is biased with the main bias term equal to $\abs{k}\Gamma(k)/n$, ignoring the small order terms arising due to estimation of $\mu$. An unbiased estimator with a divisor of $n - \abs{k}$ instead of $n$ exists, say $\Gamma^*(k)$. Nevertheless, the biased estimator is preferred for its lower mean square error (\cite{priestley1981spectral}). For a small $k$ relative to $n$, the two estimators are almost the same. However, for larger relative $k$ the variance of $\hat{\Gamma}(k)$ compensates for larger bias rendering a smaller mean square error than $\Gamma^*(k)$.

@@

{\color{blue} Fix the proof for this.}
\begin{theorem} \label{th:G-_expec} Under stationarity,
\[
   \mathbb{E} \left[\hat{\Gamma}_{G,s}(k) \right] = \left(1- \dfrac{|k|}{n}\right) \left(\Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2}\right)  + o \left(n^{-2} \right)\,,
\]
and consequently,
\[
\mathbb{E} \left[\hat{\Gamma}_{G,s}(k) - \hat{\Gamma}(k)_s \right] = \dfrac{\abs{k}}{n}\left(1 - \dfrac{1}{m}\right)\left(\dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2}\right) + o\left(n^{-2} \right)\,. 
\]
   % \[
   % \mathbb{E} \left[\hat{\Gamma}_{G}(k) \right] = \left(1- \dfrac{|k|}{n}\right)\left(\Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2}\right)  + \dfrac{2(m-1)}{m}\dfrac{\abs{k}}{n^2} \left(\Sigma - \sum_{h=0}^{n-1}\Gamma(h)\right) + o(n^{-2})
   % \]
\end{theorem}

% \begin{proof}
%  The estimator for autocovariance with divisor $n$ is suggested by \cite{parzen1961approach} for its lesser mean square error despite having greater bias than the unbiased estimator with an $n - \abs{k}$ divisor. 
%  We know that $\hat{\Gamma}_s(k)$ is asymptotically unbiased with a $o(1)$ bias term that remains fixed for any $k$. The bias term is given by \cite{priestley1981spectral} as

@@

By proposition 1 in \cite{song1995optimal} 
\begin{equation} \label{eq:song}
\Var(\bar{X}) = \dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2} + o(n^{-2})
\end{equation}

@@

{\color{blue} Below is a remark}
\begin{corollary} \label{cor:lag0_expectation}
\[
\mathbb{E} \left[\hat{\Gamma}_{G}(0) \right] = \Gamma(0) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2} + o(n^{-2})
\]

\end{corollary}

@@

% subsection autocorrelation (end)
{\color{blue}  Autocovariances will be used implictly in SVE estimators. But they are explicitly important in ACF plots via autocorrelations.

Notation in the appendix is $\Gamma^{ii}_{G,s}$
}

For any component $i$, the autocorrelation is
\[
\rho^{(i)}(k) = \dfrac{\Gamma^{(i)}(k)}{\Gamma^{(i)}(0)}\,,
\]
and is particularly useful for visualizing the serial correlation in components of the Markov chain. The typical estimate of the autocorrelation is constructed from  locally-centered autocovariance estimates which as was demonstrated in Section~\ref{sec:intro} can significantly underestimate the autocorrelation. 
\[
\hat{\rho}_{G,s}^{(i)}(k) = \dfrac{ \hat{\Gamma}^{(i)}_{G,s} (k)}{\hat{\Gamma}^{(i)}_{G,s} (0)}
\]
{\color{blue} Add more about autocorrelation. Refer to Figure 2 above and the impacts of the correlation methods. }

@@

Suppose $F$ is the target distribution on support $\chi \subseteq \mathbb{R}^d$. For $g:\chi \longrightarrow \mathbb{R}^p$ be an F-integrable function. We are interested in
%
\[
\mu_g = \mathbb{E}_F[g(X)] = \int_{\chi}g(x)F(dx)
\]

Let $\{X_{t}\}_{t \geq 1}$ denote the Harris ergodic (aperiodic, F-irreducible, and positive Harris recurrent) Markov chain having invariant distribution $F$. We will use the following Monte Carlo estimator in order to estimate $\mu$
%
\[
\hat{\mu}_s = \dfrac{1}{n}\sum_{i = 1}^{n} g(X_{si}) \xrightarrow{a.s.} \mu \textrm{ as } n \to \infty
\]

We are interested in Markov chain average and therefore, throughout this paper, we will deal with the function $g(X) = X$. To assess the reliability of our simulations, we are interested in Monte Carlo error, i.e. $\bar{\bar{X}} - \mu$. The sampling distribution for Monte Carlo error is available via Markov chain CLT (\cite{jones2004markov}) if there exists a positive definite matrix $\Sigma$ such that
%
{\color{blue} }
\[
\sqrt{n}(\bar{X}_s-\mu) \xrightarrow{d} N(0,\Sigma)
\]

 It is important to construct an estimator for $\Sigma$ which is very close to underlying reality. There are many techniques for estimating $\Sigma$ like batch means (BM), overlapping bath means (OBM), and spectral variance estimator (SVE). Due to its nice statistical properties and heavy dependence on autocovariance, we will only focus on SVE in this paper. SVE involves a truncated and weighted sum of autocovariances. We will give theoretical as well as experimental results for the version of SVE that uses our globally centered autocovariances. We call this globally-centered SVE (G-SVE).

Given m parallel Markov chains, we have $m$ i.i.d. samples of $\bar{X}_s$, $s \in \{1,..., m\}$. The Markov Chain standard error (MCSE) is given by $\Sigma/mn$. We are interested in consistent estimates of $\Sigma$ for two reasons. Firstly, to construct asymptotically valid confidence intervals and secondly, to determine when to stop the simulations. For this purpose, we use the class of Multivariate Spectral Variance Estimators (MSVE). The most common practice would be to combine the spectral variance estimates from $m$ chains by averaging over them. We call this technique average spectral variance estimator and denote it by $\hat{\Sigma}_A$). We propose using a globally-centered spectral variance estimator (G-SVE) wherein we use G-ACF to estimate the autocovariance. Strong consistency for MSVE has been addressed by \cite{vats:fleg:jon:2018}. Bias and variance calculations for a variant of SVE with known mean is done by \cite{hannan2009multiple}. In this paper, we will address the asymptotic properties (strong consistency), bias and variance calculations for G-SVE.

@@

% Consider $m$ Markov chains where $\hat{\Gamma}_s(k)$, $\hat{\Gamma}_{G,s}(k)$ and $\hat{\Sigma}_{S, s}$ are the lag-$k$ empirical ACF, G-ACF, and SVE respectively for $s^{th}$ Markov chain, $s\in {1,...,m}$. 
Consider the globally-centered SV estimator (G-SVE) for the $s$th chain using G-ACF:
%
\begin{align*}
    \hat{\Sigma}_{G,s} &= \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Gamma}_{G,s}(k)\,,
\end{align*}
and the combined G-SVE is
\[
\hat{\Sigma}_{G} =  \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Sigma}_{G,s}\,.
\]

@@

\begin{ass}[Strong Invariance Principle (SIP)] \label{ass:sip}
    We assume the SIP holds for the Markov chain. Here $\{X_t\}_{t\geq 1}$ is the stochastic process that follows SIP with mean $\mu$. Let $\|.\|$ denote the Euclidean norm. Through SIP, there exists a $p \times p$ lower triangular matrix L, an increasing function $\psi$ on the integers, a finite random variable D, and a  sufficiently rich probability space such that, with probability 1, \\
  $$\left\|\sum_{t=1}^{n}X_t - n\mu - LB(n)\right\| < D\psi(n)$$

@@

{\color{blue} Don't make this an assumption. Say this explicitly in Theorem 2. }

\begin{ass} \label{ass:sve_consis} We assume that all the conditions given by \cite{vats:fleg:jon:2018} for strong consistency of the spectral variance estimator hold true. As a consequence, $\hat{\Sigma}_{S,s} \xrightarrow{a.s.} \Sigma$ as $n \to \infty$ for all Markov chains $s$.
% \[
% \hat{\Sigma}_{S,s} \xrightarrow{a.s.} \Sigma \textrm{ as } n \to \infty \;\; \forall s \in \{1,..., m\}
% \]
\end{ass}

\begin{theorem}
\label{th:consistency}
 Let the Assumptions \ref{ass:sip},\ref{ass:sve_consis} hold and $n^{-1}{b_n \log \log n} \to 0 \textrm{ as } n \to \infty$, then $\hat{\Sigma}_{G} \to \Sigma$ with probability 1, as $n \to \infty$.
\end{theorem}

@@

{\color{blue}: Make the following an assumption. Is $2\pi$ needed? Is $W_n$ always finite? }
We are interested in the finding the bias for G-SVE estimator as a function of $n$. The proof below applies to all the important range of lag windows $w(x)$; where $w(x)$ is a continuous, even fnction with $w(0)=1, \abs{w(x)} < c$, and $\int_{-\infty}^{\infty}w^2(x)dx < \infty$. Define
%
\[
W_n = \dfrac{1}{2\pi}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right).
\]

We have $W_n < \infty$ for all the important lag windows.

\begin{ass} \label{ass:bias}
    Let $\Gamma_s(k)$ be the lag-k autocovariance for $s^{th}$ Markov chain and $w(x)$ be the lag window. We assume that there exists a $q \geq 0$ such that for all $s$
    \begin{enumerate} [a.]
        \item $\sum\limits_{-\infty}^{\infty}\abs{k}^q \|\Gamma_s(k)\| < \infty$
        \item $\lim _{x \to 0}\dfrac{1 - w(x)}{\abs{x}^q} = k_q < \infty$
        \item $\dfrac{b_n^q}{n} \to 0$ as $n \to \infty$
    \end{enumerate}

@@

{\color{blue} Under what assumptions?}
\begin{theorem}\label{th:G-SVE_bias}
The limiting bias of replicated spectral variance is given by:
\[
 \lim_{n \to \infty}b_n^q\mathbb{E} \left[\hat{\Sigma}_{G} - \Sigma \right] = -k_q\Phi^{(q)}
 \]
\end{theorem}

@@

{\color{blue} Don't make this a separate assumption. Include in the statement of the theorem}
\begin{ass} \label{ass:variance_cal}
Let $\{X_{st}\}_{t \geq 1}$ be the $s^{th}$ Markov chain for which SIP holds. 
\begin{enumerate}[a.]
    \item $\E[D^4] < \infty$ where D is a finite random variable associated with the SIP
    \item $\E[X^4] < \infty$ where $\{X\}$ is the stochastic process
\end{enumerate}
\end{ass}

\begin{theorem} \label{th:G-SVE_variance}
 If Assumption \ref{ass:variance_cal} holds, $b_n^{-1}{n}\Var \left(\hat{\Sigma}_{G}^{ij} \right) = [\Sigma_{ii}\Sigma_{jj} + \Sigma_{ij}^2]\int_{-\infty}^{\infty}w(x)^2dx  + o(1)$.
\end{theorem}

@@

>>>>

The power of the modern personal computer has made it easy to run parallel Markov chain Monte Carlo (MCMC) implementations. This is particularly useful for slow mixing or multi-modal target distributions, where the starting points of the chains are spread over the state-space in order to more accurately capture characteristics of the target distribution. Output from $m$ parallel chains are then summarized, visually and quantitatively, to assess the empirical mixing properties of the chains and the quality of Monte Carlo estimators.

A key quantity of interest that drives MCMC output analysis is the autocovariance function (ACvF). From their use in autocorrelation plots to assessing Monte Carlo variability of estimates, to determining when to stop MCMC simulation, autocovariances drive many visual and quantitative inferences users make from MCMC output. However, tools on estimating ACvF are largely constructed for output from one MCMC chain.

% Markov chain Monte Carlo (MCMC) is a sampling method used to obtain serially correlated realizations from a desired probability distribution when i.i.d. sampling is difficult. Due to this serial correlation in samples, a larger number of MCMC samples are required to make inferences about the target distribution as compared to the independent ones. Therefore, it becomes imperative for an MCMC practitioner to have an idea about the underlying dependence structure in the Markov chains. Autocovariance function (ACF) characterizes this dependence at a particular lag $k$.

Let $F$ be the target distribution with mean $\mu$ defined on the set $\X \subseteq \mathbb{R}^p$ equipped with a countably generated Borel $\sigma$-field $\mathcal{B}(\X)$. Let $\{X_1, X_2, ..., X_n\}$ be an $F$-Harris ergodic (aperiodic, $F$-irreducible, and positive Harris recurrent) Markov chain. 
% If $\bar{X} = n^{-1} \sum_{t=1}^{n}X_t$ , then with probability 1, $\bar{X} \to \mu$ as $n \to \infty$. \\

For a stationary Markov chain, the autocovariance function at lag $k$ depends only on $k$ and is defined as 
%
\[
    \Gamma(k) = \Cov(X_t, X_{t+k})= \mathbb{E}[(X_t - \mu)(X_{t+k} - \mu)^T].
\]

A popular sample autocovariance matrix at lag-$k$ is given by
%
\begin{equation} \label{eq:empirical_acf}
    \hat{\Gamma}(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}}(X_t - \bar{X})(X_{(t+\abs{k})} - \bar{X})^T.
\end{equation}

It is a known fact that most of the Markov chains obtained from practical MCMC algorithms are not stationary. Regardless, Equation~\ref{eq:empirical_acf} provides the empirical estimator for ACF at a particular lag for the Markov chain with any initial distribution (\cite{geyer2011introduction}).\\

Calculation of exact ACF has been done for simple stationary stochastic models like autoregressive (AR) models (under assumption of stationarity), moving average (MA) (\cite{quenouille1947notes}), and autoregressive-moving-average (ARMA) (\cite{box2015time}) models. However, for most of the complex sampling algorithms (including MCMC), we rely on sample estimators for ACF. Unfortunately, the estimator given in Equation~\ref{eq:empirical_acf} has been observed to have a some drawbacks. Firstly, it is not immune to outliers (see \cite{ma2000highly}) and secondly, it is usually non-informative in case of slowly mixing Markov chains. We address the second issue in this paper by using multiple representative samples of the target distribution via parallel Markov chains. This approach is especially relevant to MCMC wherein multiple chain sampling is computationally possible and statistically desirable for making informative inferences. Our goal is to pool in the information from all the Markov chains while estimating the autocovariances. \\

 A commonly faced problem with the MCMC models is the slow mixing property. Especially in case of multi modal target distributions, it has been observed that the Markov chains tend to get \textit{stuck} in a particular mode for a long time. In such situations, single chain empirical ACF will severely underestimate the autocovariance. We propose an autocovariance estimator called globally-centered autocovariance function (G-ACF) with a simple fix in centering which has been observed to show drastically better results in many cases. We also prove that our estimator compensates the negative bias of the empirical ACF estimator to give estimates that are closer to reality. We will further study the impact of G-ACF in calculation of two important applications of ACF (1) asymptotic variance of Markov chain average and (2) effective sample size (ESS).\\

Let $g: \chi \to \mathbb{R}^d$ be an F-integrable function and suppose we are interested in estimating the expected value $\mu_g = \mathbb{E}_F[g(X)]$. We know that $\sum_{t=1}^{n}g(X_t)/n$ converges to $\mu_g$ as $n \to \infty$ almost surely. MCMC ensures asymptotic convergence, however this is not practically very useful where the number of samples is finite and the Markov chain mixes slowly. One of the important criteria to determine the quality of estimation is the Monte Carlo aymptotic covariance matrix which is given by an infinite summation of autocovariance matrices for all lags. For a slowly mixing Markov chain, underestimated autocovariances will deliver underestimated asymptotic variance. There are many techniques for estimating it like batch means (BM), overlapping bath means (OBM), and spectral variance estimator (SVE). Due to its nice statistical properties, we will only focus on SVE in this paper. SVE involves a truncated and weighted sum of autocovariances. We propose a version of SVE that uses our globally centered autocovariances. We call this globally-centered SVE (G-SVE).\\

A more powerful estimator of Monte Carlo asymptotic covariance matrix can enable us to estimate the effective sample size for the Markov chain which is closer to reality. There are many criteria for terminating simulation like fixed-time rule, sequencial fixed-volume rule (see \cite{glynn1992asymptotic}), and using Gelman Rubin diagnostics (\cite{gelman1992inference}). The relative fixed-volume rule has been thoroughly discussed for MCMC by \cite{flegal2015relative} and \cite{gong2016practical}. We will use method of effective sample size (ESS) by \cite{vats2019multivariate} for terminating the simulations. The quality of estimation of asymptotic covariance matrix is critical in estimation of the ESS. We will successfully show that the G-SVE estimates the truth better and prevents the premature stopping of Markov chain.\\

The rest of the paper is arranged as follows. In Section~\ref{sec:G-ACF} we introduce the new sample autocovariance estimator and calculate its bias for finite samples. We observe that the bias term for G-ACF is similar to the bias of empirical ACF except for a few small order end-effect terms that vanish as $n \to \infty$. In Section~\ref{sec:variance_est}, we introduce the globally-centered SVE and give a proof of strong consistency, bias, and variance calculations. We will elaborate on the estimator for ESS using G-SVE in Section~\ref{sec:ess}. Section~\ref{sec:examples} gives an experimental study on three different examples for the proposed estimators. All the important proofs are given in the Appendix.

@@

We demonstrate the striking difference in the ACF plots through mixture of Gaussian target density using a random-walk Metropolis-Hastings sampler. The target density has two well separated and non-equal modes. The target distribution $F$ defined on $\mathbb{R}$ is given by
\[
F(x) = 0.7\:\textrm{N}(-5, 1) + 0.3\: \textrm{N}(5, 0.5)
\]
Two Markov chains are started close to each mode. It can be seen in Figure \ref{fig:gaussian-trace} that for the first $10^4$ samples, the chains do not communicate with the other mode. Whereas for a sample size of $10^5$, the chains have jumped modes and therefore, a single chain can give reliable information about the distribution. We wish to make two points here: firstly, for small sample size, we observe that ACF gives extremely misleading estimates of autocovariance whereas G-ACF is very close to reality (Figure \ref{subfig:gauss-acf-1e4}). Secondly, for a large sample size, the estimates from G-ACF as well as ACF are equivalent. Figure \ref{subfig:gauss-trace_1e5} depicts the scenario when convergence is achieved. This shows that G-ACF is at least as good as ACF; hence safe to use in any setting.

@@

 Suppose we have $m$ replications of Markov chains for same target distribution and $\{X_{st}; \; t \in \mathbb{Z}\}$ denote the $s^{th}$ Markov chain. Let the sample mean of $s^{th}$ Markov chain be denoted by $\bar{X}_s$ and the global mean by $\bar{\bar{X}} = \sum_{s = 1}^{m}\bar{X}_s/m$. The global mean provides a better estimate for the expectation value when the number of simulation per chain are not sufficient for the convergence to kick in. The globally-centered autocovariance function (G-ACF) for the $s^{th}$ Markov chain is given by:
%
\[
\hat{\Gamma}_{G,s}(k) = \dfrac{1}{n} \sum_{t=1}^{n-k}(X_{st}-\bar{\bar{X}})(X_{s(t+k)}-\bar{\bar{X}})^T,
\]
and the averaged globally-centered ACF over all $m$ chains is given by 
\[
\hat{\Gamma}_G(k) = \dfrac{1}{m}\sum_{s=1}^m \hat{\Gamma}_{G,s}(k)\;.
\]
If the starting points of these parallel Markov chains are sufficiently dispersed, G-ACF provides more realistic estimation of lag covariances. \cite{priestley1981spectral} shows that the empirical estimator in Equation \ref{eq:empirical_acf} is biased with the main bias term equal to $\abs{k}\Gamma(k)/n$, ignoring the small order terms arising due to estimation of $\mu$. Though, an unbiased estimator with a divisor of $n - \abs{k}$ instead of $n$ exists, the biased estimator is preferred for its lower mean square error for larger relative $k$  (\cite{priestley1981spectral})

@@

{\color{blue} Fix the proof for this.}
\begin{theorem} \label{th:G-_expec} Under stationarity,
\[
   \mathbb{E} \left[\hat{\Gamma}_{G,s}(k) \right] = \left(1- \dfrac{|k|}{n}\right) \left(\Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2}\right)  + o \left(n^{-2} \right)\,,
\]
and consequently,
\[
\mathbb{E} \left[\hat{\Gamma}_{G,s}(k) - \hat{\Gamma}(k)_s \right] = -\dfrac{\abs{k}}{n}\Gamma(k) - \dfrac{1}{m}\left(1- \dfrac{|k|}{n}\right)\left(\dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2}\right) + o\left(n^{-2} \right)\,. 
\]
   % \[
   % \mathbb{E} \left[\hat{\Gamma}_{G}(k) \right] = \left(1- \dfrac{|k|}{n}\right)\left(\Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2}\right)  + \dfrac{2(m-1)}{m}\dfrac{\abs{k}}{n^2} \left(\Sigma - \sum_{h=0}^{n-1}\Gamma(h)\right) + o(n^{-2})
   % \]
\end{theorem}

% \begin{proof}

%  We know that $\hat{\Gamma}_s(k)$ is asymptotically unbiased with a $o(1)$ bias term that remains fixed for any $k$. The bias term is given by \cite{priestley1981spectral} as

@@

@@

\begin{remark} \label{rem:lag0_expectation}
\[
\mathbb{E} \left[\hat{\Gamma}_{G}(0) \right] = \Gamma(0) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2} + o(n^{-2})
\]

\end{remark}

@@

% subsection autocorrelation (end)
%{\color{blue}  Autocovariances will be used implictly in SVE estimators. But they are explicitly important in ACF plots via autocorrelations.

%Notation in the appendix is $\Gamma^{ii}_{G,s}$
%}
Lag-0 autocovariance are particularly useful in calculating the autocorrelations.The variance of $i^{th}$ component given by $\Gamma^{ii}(0)$. For any component $i$, the autocorrelation is
\[
\rho^{ii}(k) = \dfrac{\Gamma^{ii}(k)}{\Gamma^{ii}(0)}\,,
\]
and is instrumental in visualizing the serial correlation in components of the Markov chain. The typical estimate of the autocorrelation is constructed from  locally-centered autocovariance estimates which as was demonstrated in Section~\ref{sec:intro} can significantly underestimate the autocorrelation. Instead, we calculate the globally-centered autocorrelation using the formula
\[
\hat{\rho}_{G,s}^{(i)}(k) = \dfrac{ \hat{\Gamma}^{(i)}_{G,s} (k)}{\hat{\Gamma}^{(i)}_{G,s} (0)}.
\]
Despite using similar centering in both numerator and denominator, the globally-centered autocorrelation is far greater than locally-centered autocorrelation for each component as is evident from Figure~\ref{fig:gauss:acf}.

@@

Let $g:\chi \longrightarrow \mathbb{R}^d$ be an F-integrable function. We are interested in the expected value
%
\[
\mu_g = \mathbb{E}_F[g(X)] = \int_{\chi}g(x)F(dx)\;.
\]

Let $Y$ denote the random variable $g(X)$ where $X \sim F$. Since we have $m$ chains of $n$ samples each, let us make following notations: (1) $Y_{st} = g(X_{st})$ for $t \in \{1, 2, ..., n\}$ and $s \in \{1, 2, ..., m\}$; (2) $\bar{Y}_s = \sum_{t=1}^{n}g(X_{st})/n$; and (3) $\bar{\bar{Y}} = \sum_{s=1}^{m}\bar{Y}_s/m$. To assess the reliability of our simulations, we are interested in Monte Carlo error, i.e. $\bar{Y_s} - \mu$ for the $s^{th}$ chain. The sampling distribution for Monte Carlo error is available via Markov chain CLT (\cite{jones2004markov}) if there exists a positive definite matrix $\Sigma$ such that
%
{\color{blue} }
\[
\sqrt{n}(\bar{X}_s-\mu) \xrightarrow{d} N(0,\Sigma)
\]

 Given m parallel Markov chains, we have $m$ i.i.d. samples of $\bar{X}_s$, $s \in \{1,..., m\}$. The Markov Chain standard error (MCSE) is given by $\Sigma/mn$. We are interested in consistent estimates of $\Sigma$ for two reasons. Firstly, to construct asymptotically valid confidence intervals and secondly, to determine when to stop the simulations. For this purpose, we use the class of Multivariate Spectral Variance Estimators (MSVE). The most common practice would be to combine the spectral variance estimates from $m$ chains by averaging over them. We call this technique average spectral variance estimator (A-SVE) and denote it by $\hat{\Sigma}_A$. Let $\hat{\Sigma}_{SV,s}$ denote the SVE for the $s^{th}$ chain, then the A-SVE is given by

\[
\hat{\Sigma}_A = \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Sigma}_{SV,s}\;.
\]
 We propose a globally-centered spectral variance estimator (G-SVE) wherein we use G-ACF to estimate the autocovariance. Strong consistency for MSVE has been addressed by \cite{vats:fleg:jon:2018}. Bias and variance calculations for a variant of SVE with known mean is done by \cite{hannan2009multiple}. In this paper, we will address the asymptotic properties (strong consistency), bias and variance calculations for G-SVE.

@@

% Consider $m$ Markov chains where $\hat{\Gamma}_s(k)$, $\hat{\Gamma}_{G,s}(k)$ and $\hat{\Sigma}_{S, s}$ are the lag-$k$ empirical ACF, G-ACF, and SVE respectively for $s^{th}$ Markov chain, $s\in {1,...,m}$. 
Consider the globally-centered SV estimator (G-SVE) for the $s$th chain using G-ACF:
%
\begin{align*}
    \hat{\Sigma}_{G,s} &= \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Gamma}_{G,s}(k)\,,
\end{align*}
and the combined G-SVE is
\[
\hat{\Sigma}_{G} =  \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Sigma}_{G,s}\,.
\]
Define
%
\[
W_n = \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right).
\]
We will consider a range of lag windows $w(x)$; where $w(x)$ is a continuous, even function with $w(0)=1, \abs{w(x)} < c$, $\int_{-\infty}^{\infty}w^2(x)dx < \infty$, and $W_n < \infty$.

@@

\begin{ass}[Strong Invariance Principle (SIP)] \label{ass:sip}
    We assume the SIP holds for the Markov chains $\{Y_{st}\}_{t \geq 1} \; \forall s$. Here $\{Y_t\}_{t\geq 1}$ is the stochastic process that follows SIP with mean $\mu_g$. Let $\|.\|$ denote the Euclidean norm. Through SIP, there exists a $p \times p$ lower triangular matrix L, an increasing function $\psi$ on the integers, a finite random variable D, and a  sufficiently rich probability space such that, with probability 1, \\
  $$\left\|\sum_{t=1}^{n}Y_t - n\mu_g - LB(n)\right\| < D\psi(n)$$

@@

\begin{theorem}
\label{th:consistency}
 Let the Assumptions \ref{ass:sip} hold, $\hat{\Sigma}_{SV,s} \xrightarrow{a.s.} \Sigma \; \forall s$, and $n^{-1}{b_n \log \log n} \to 0 \textrm{ as } n \to \infty$, then $\hat{\Sigma}_{G} \to \Sigma$ with probability 1, as $n \to \infty$.
\end{theorem}

@@

We are interested in the finding the bias for G-SVE estimator as a function of $n$. 
\begin{ass} \label{ass:bias}
    Let $\Gamma_s(k)$ be the lag-k autocovariance for $s^{th}$ Markov chain and $w(x)$ be the lag window. We assume that there exists a $q \geq 0$ such that for all $s$
    \begin{enumerate} [a.]
        \item $\sum\limits_{-\infty}^{\infty}\abs{k}^q \|\Gamma_s(k)\| < \infty$
        \item $\lim _{x \to 0}\dfrac{1 - w(x)}{\abs{x}^q} = k_q < \infty$
        \item $\dfrac{b_n^q}{n} \to 0$ as $n \to \infty$
    \end{enumerate}

@@

\begin{theorem}\label{th:G-SVE_bias}
Let the Assumption~\ref{ass:bias} hold, then the limiting bias of G-SVE is given by:
\[
 \lim_{n \to \infty}b_n^q\mathbb{E} \left[\hat{\Sigma}_{G} - \Sigma \right] = -k_q\Phi^{(q)}
 \]
\end{theorem}

@@

\begin{theorem} \label{th:G-SVE_variance}
 Let Assumption~\ref{ass:sip} holds, $\E[D^4] < \infty$ where D is a finite random variable associated with the SIP, and  $\E[Y^4] < \infty$, then $b_n^{-1}{n}\Var \left(\hat{\Sigma}_{G}^{ij} \right) = [\Sigma_{ii}\Sigma_{jj} + \Sigma_{ij}^2]\int_{-\infty}^{\infty}w(x)^2dx  + o(1)$.
\end{theorem}

@@

% \begin{proof}
